---
title: "AKurtin SSF Homework"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

```{r Load Packages, include=FALSE}
packages <- c("sf","terra","lubridate", "tidyverse","ggplot2","mapview","maptools","leaflet","xtable","broom","stars","magrittr","cowplot", "tmap","suncalc", "survival", "amt", "glmmTMB", "TMB","coxme")

#function to install and load required packages
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

# SSF Homework

## Part 1: Data Preparation

To start, I'll be reading in the wolf GPS data. It looks like summer has data on less packs, but I'm interested in examining the effects of the covariates I looked at in the RSF in the context of summer. I am interested to see what their habitat selection looks like when they aren't as constrained by ice and snow.

```{r Load in Wolf Data}
wolfGPS <- read.csv("Data/wolf_data.csv")
# Take a look at what this GPS data looks like
ggplot(wolfGPS, aes(x_coord1, y_coord1, colour = pack)) +
  geom_point()+
  labs(title ="All Wolf GPS Locations")

# Pick a season:
wolf_summer <- wolfGPS %>% filter(season=="summer")
# unique(wolf_summer$wolfuid)
ggplot(wolf_summer, aes(x_coord1, y_coord1, colour = pack)) +
  geom_point() +
  labs(title ="Wolf Summer GPS Locations")
# Summer only gives us ranch and red deer

wolf_winter <- wolfGPS %>% filter(season=="winter")
ggplot(wolf_winter, aes(x_coord1, y_coord1, colour = pack)) +
  geom_point() +
  labs(title ="Wolf Winter GPS Locations")
# winter gives us more points
```

Next, I'll clean up the data and create a shapefile from it. 
```{r Clean and Convert to Shapefile}
# convert wolf data into a shapefile
# Select out only columns that have a lat and long 
wolf2<-wolf_summer[complete.cases(wolf_summer[10:11]),]

# Unite date and time into a column
wolf_summer2 <- wolf2 %>% unite(datetime,c(date,time),sep=" ")

# Reformat the date and time column into class datetime 
wolf_summer2 <- wolf_summer2 %>% mutate(datetime_new=as.POSIXct(datetime,format="%m/%d/%y %H:%M"))

# remove the NAs in the datetime_new column
wolf_summer2<-wolf_summer2[complete.cases(wolf_summer2[24]),]
# Check to make sure there are no NA's
wolf_summer2 %>% filter(is.na(xcoord)==TRUE)
wolf_summer2 %>% filter(is.na(ycoord)==TRUE)
wolf_summer2 %>% filter(is.na(datetime_new)==TRUE)

# for future plotting, make the wolf data into a shapefile
# define the lat long as epsg 4326
wolf_sf <- st_as_sf(wolf_summer2,
                      coords = c("long","lat"),
                      crs = "EPSG:4326")
```

Next, I'll load in the spatial data. The two covariates I am interested in looking at in this analysis is conifer land cover and elevation.

```{r Load in Raster Data}
# pull the landcover data in from lab 3
lancov <- rast(".\\Data\\landcover16.tif")
# Check the crs for this raster
crs(lancov) # looks like it is in crs 26911, so let's change it to match our wolf shapefile
# match the projection
lancov2 <- lancov %>% terra::project("EPSG:4326", method = "near")
# looking at conifer is 1,2, or 3
# select the levels of lancov that are 1,2, or 3
conif <- lancov2$HABITATTYPE %in% c("Open Conifer","Moderate Conifer","Closed Conifer")
plot(conif)
#convert to a raster for future use
conif_r <- as(conif, "Raster")
# The name of this raster is undefined, so let's name it
names(conif_r) <- "conifer"

## Test how data looks
plot(conif, main = "Conifer Land Cover and Wolf Points")
plot(wolf_sf, add=TRUE) # two points are off here but overall it looks fine


# Now let's pull in the elevation data
elevation <- rast(".\\Data\\Elevation2.tif")
# check the projection
crs(elevation) #26911
# reproject to match the others
elev2 <- elevation %>% terra::project("EPSG:4326", method = "near")
# convert to a raster for future use
elev2_r <- as(elev2, "Raster")

## Test how data looks
plot(elev2, main = "Elevation and Wolf Points")
plot(wolf_sf$geometry, add=TRUE) # This looks good now
```


To fit my SSF, I need to take the wolf GPS points and make track data.

```{r Set up Track Data}
# nest the data by the wolf ID
wolf_summer3 <- wolf_summer2 %>% nest(data = c(-wolfuid,-pack))

# Make the track data with a function
track_all <- wolf_summer3 %>%
 # make the track
 mutate(trk = map(data, function(d) {
   #transform from geographic to projected coord system 
    amt::make_track(tbl=d, .x=long, .y=lat, .t=datetime_new, crs = 4326) %>% # changed from tbl=d, .x=xcoord, .y=ycoord, .t=datetime_new
      amt::transform_coords(st_crs(conif))}))


#summarize sampling rates by individuals and then unnest 
samp_rates <- track_all %>% mutate(sr = lapply(trk, summarize_sampling_rate)) %>%
 dplyr::select(wolfuid, sr) %>% unnest(cols = c(sr))
samp_rates

```

Looking at the sampling rates, I see that the highest median fix rate is 6 hours, so I'll be resampling them all to 6 hours. 

Now, I'll extract covariates to make an SSF dataframe. 

```{r}
# Make an SSF dataframe
track_resamp_all <- track_all %>%
   mutate(steps = map(trk, function(x) {
     x %>% amt::track_resample(rate = minutes(360), tolerance = minutes(10)) %>% # resample the tracks for each individual
      amt::filter_min_n_burst() %>% # filter them into bursts
       amt::steps_by_burst() %>% amt::random_steps() %>% # filter them into bursts again
       amt::extract_covariates(covariates = elev2, where = "both") %>% # this does both start and end | names the column the same as the column and adds on where 
       mutate(unique_step = paste(burst_,step_id_,sep="_")) %>% 
#       mutate(elev_start = factor(Elevation2_start)) %>% 
#       mutate(elev_end = factor(Elevation2_end)) %>% # filter them into bursts again
       amt::extract_covariates(covariates = conif, where = "both") %>% # this does both start and end | names the column the same as the column and adds on where 
       mutate(conifer_start = factor(HABITATTYPE_start)) %>% 
       mutate(conifer_end = factor(HABITATTYPE_end)) 
     }))

# Unnest this data
tracks_unnest <- track_resamp_all %>% 
  dplyr::select(wolfuid,pack, steps) %>% 
  unnest()

# Clean up the data
tracks_unnest <- tracks_unnest %>% rename(elev_start = Elevation2_start, elev_end = Elevation2_end) %>% select(-c(HABITATTYPE_start,HABITATTYPE_end))

# Make a column for pack as a factor
tracks_unnest <- tracks_unnest %>% mutate(pack_fac=factor(pack))
```

## Part 2: Statistical Analysis 

First, let's do some exploratory analyses of the effects of the covariates on step length to see if we need to include an interaction. 

```{r Exploratory Analysis}
# Look at step length and turning angle distributions
## sl_ for step length and ta_ for turning angle, time end is t2_

# for making plots:
tracks_plotting <- tracks_unnest %>% filter(is.na(conifer_end)==FALSE) %>% filter(is.na(elev_end)==FALSE)

# Relationship between step length and conifer land cover
# make a bar plot of sl_ for conifer_end = TRUE and FALSE
ggplot(data = tracks_plotting, mapping = aes(x = conifer_end, y = sl_)) + 
  geom_boxplot() + facet_wrap(~case_)
# doesn't look like there's a difference

# Relationship between elevation and land cover
ggplot(data = tracks_plotting, mapping = aes(x = elev_end, y = sl_)) + 
  geom_point()+ facet_wrap(~case_)
# slight increase at the end but no clear trend

# Test for interactions with step length with covariates
ggplot(data = tracks_plotting, mapping = aes(x = elev_end, y = sl_, color = conifer_end)) + 
  geom_point()+ facet_wrap(~case_)

 
```

We also don't really see a trend here in the step length of our used points in relation to our covariates, so we don't need to include an interaction. Now, let's fit some models to our data. 

**1. Naive GLM**

```{r}
# Fit a naive GLM
nGLM <- glm(case_~ I(conifer_end)+elev_end, data=tracks_unnest,family=binomial(link="logit"))
# Take a look at the exponentiated coefficients
coefs_nGLM <- exp(nGLM$coefficients)

```

Here, we see that there is a very weak (litte to no effect) of elevation on selection, and there is a negative effect of conifer land cover on selection. 

**2. Step Selection Function**

```{r SSF}
# Fit an SSF Model
# hypothesis that habitat selection varies as a function of conifer presence, elevation, and the interaction of the conifer and elevation
ssf_mod <- tracks_unnest %>% amt::fit_issf(case_ ~ conifer_end + elev_end + conifer_end:elev_end + strata(unique_step))
summary(ssf_mod)
ssf_sum <- summary(ssf_mod$model)$coefficients

# Look at the outputs 
ssf_sum
AIC(ssf_mod)
```

Here, we see that again there is little to no effect of elevation and no effect of the interaction between elevation and conifer. The effect of conifer is negative and is stronger than elevation.

**3. Clogit**

```{r Naive Clogit}
# Fit a clogit model
clogit1 <- clogit(case_ ~ I(conifer_end) +(elev_end) + strata(unique_step),data = tracks_unnest)
#clogit1 <- clogit(case_ ~ I(landuseName=="Developed Open") + I(landuseName=="Developed Other") +I(landuseName=="Natural")+I(landuseName=="Crops") + strata(stratum), data = fisher6)
summary(clogit1)
exp(coef(clogit1))
AIC(clogit1)
```

Here, we seee that there is little to no effect of elevation on selection and a negative effect of conifer land cover on selection.

**4. Coxme**

```{r coxme model}
# Create a time variable
tracks_unnest$time_ <- ifelse(tracks_unnest$case_ == 0, 2, 1)   #2 for control, 1 for case
# Fit a coxme model with random intercepts and random slopes 
clogitM1<- coxme(Surv(time_,case_) ~ conifer_end + elev_end + conifer_end:elev_end  + (1|pack_fac) + (0+pack_fac) , data=tracks_unnest)
AIC(clogitM1)
summary(clogitM1)
```

Here, we see that there is little to no effect of elevation and the conifer:elevation interaction on selection. Conifer landcover has a negative effect on selection, and the pack identify had a positive effect of selection.


**5. iSSF with glmmTMB**

```{r glmmTMB}
# Fit glmmTMB model with random effect of step ID and random slope of pack on conifer 
TMBm1 <- glmmTMB(case_~ I(conifer_end) + elev_end + (1|step_id_) + (0 + I(conifer_end)|pack_fac), family = poisson, data = tracks_unnest, doFit = FALSE)

TMBm1$parameters$theta[1] <-log(1e3) # set theta one (the 0 in the 0 + I(landuse_end)|id)
TMBm1$mapArg <-list(theta=factor(c(NA, 1:3)))
glmm.TMB.random <- glmmTMB::fitTMB(TMBm1) # fit your model (from above) using these arguments
summary(glmm.TMB.random)
# Look at the coefficients
coefTMB <- fixef(glmm.TMB.random)
```

Here, we see that elevation has little to no effect on selection and conifer land cover has a negative effect on selection.


# Part 3: Inference

In this analysis, we saw that when we compare the ssf model estimated a naive clogit, coxme, and glmmTMB, each model gave us the same interpretation, which is that elevation has little to no effect on selection but conifer land cover has a negative effect. Therefore, we can conclude that it is not necessary to incorporate the movement model to determine the effect of only elevation and conifer land cover on habitat selection, and employing a simpler model will give us the same interpreation. 

