
---
title: "SSF Homework"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SSF Homework

Due April 13th in class. 

Conduct an SSF model for wolves in the Cascade, Red Deer, Wildhorse, Ranch and Bow Valley wolf packs for some covariates that we have used this semester such as one continuous variable (elevation, slope) and 1 categorical variable (open, closed landcover types).  Test for interactions with step length with covariates. Focus just on one season, and just 1 level of random effects at the pack level including intercepts and coefficients as recommended by Muff et al. (2020).  Finally, compare your ssf model estimated with a naive clogit, coxme, and glmmTMB.  Optional (can't do this year bc of issues with the packages): Try your hand at mapping the iSSF using the movement kernel, transient UD, etc.  That might take a long time though.

No format for how to turn it in. Just write up the methods, a markdown document would be great. 
Just pick a covariates like landcover or elevation
Include step lengths and time of day (even though it's buggy) and play around with it/compare models. 
To compare models, just look at the coefficient table (last part of what we did today) (can't use AIC for these things). Don't worry about validation.

```{r}
# use the banff raster stacks from other labs 
# pull this from the RSF project
# pick one season: Summer
# one continuous variable: elevation
# one categorical variable: habitat type
# use pack as the hierarchy
# play with a couple of covariates 
# play with sampling rate
# use the set of code where you build the SSF all in one pass

# styleR

```

```{r Load Packages}
packages <- c("sf","terra","lubridate", "tidyverse","ggplot2","mapview","maptools","leaflet","xtable","broom","stars","magrittr","cowplot", "tmap","suncalc", "survival", "amt", "glmmTMB", "TMB","coxme")

#function to install and load required packages
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

# Part 1: Data Preparation

```{r Load in Wolf Data and Clean It}
wolfGPS <- read.csv("Data/wolf_data.csv")
# Take a look at what this GPS data looks like
ggplot(wolfGPS, aes(x_coord1, y_coord1, colour = pack)) +geom_point()

# Pick a season:
wolf_summer <- wolfGPS %>% filter(season=="summer")
# unique(wolf_summer$wolfuid)
#dev.off()
#ggplot(wolf_summer, aes(x_coord1, y_coord1, colour = pack)) +geom_point()
# Summer only gives us ranch and red deer
wolf_winter <- wolfGPS %>% filter(season=="winter")
#dev.off()
#ggplot(wolf_winter, aes(x_coord1, y_coord1, colour = pack)) +geom_point()
# winter gives us more points
```

Winter has more points and more packs, but summer is more interesting because they aren't as limited to where they can go

```{r Convert Wolf Data to Shapefile}
# convert wolf data into a shapefile
# Select out only columns that have a lat and long 
wolf2<-wolf_summer[complete.cases(wolf_summer[10:11]),]
#unique(wolf2$wolfuid)
# Unite date and time into a column
wolf_summer2 <- wolf2 %>% unite(datetime,c(date,time),sep=" ")

# Reformat the date and time column into class datetime 
wolf_summer2 <- wolf_summer2 %>% mutate(datetime_new=as.POSIXct(datetime,format="%m/%d/%y %H:%M"))
# format that you specify for what it's coming in as

# remove the NAs in the datetime_new column
wolf_summer2<-wolf_summer2[complete.cases(wolf_summer2[24]),]
# Check to make sure there are no NA's
wolf_summer2 %>% filter(is.na(xcoord)==TRUE)
wolf_summer2 %>% filter(is.na(ycoord)==TRUE)
wolf_summer2 %>% filter(is.na(datetime_new)==TRUE)

# for future plotting, make the wolf data into a shapefile
# define the lat long as epsg 4326
wolf_sf <- st_as_sf(wolf_summer2,
                      coords = c("long","lat"),
                      crs = "EPSG:4326")
```


```{r Load in Raster Data}
# pull the landcover data in from lab 3
lancov <- rast(".\\Data\\landcover16.tif")
# Check the crs for this raster
crs(lancov) # looks like it is in crs 26911, so let's change it to match our wolf shapefile
# match the projection
lancov2 <- lancov %>% terra::project("EPSG:4326", method = "near")
# looking at conifer is 1,2, or 3
# select the levels of lancov that are 1,2, or 3
conif <- lancov2$HABITATTYPE %in% c("Open Conifer","Moderate Conifer","Closed Conifer")
plot(conif)
#convert to a raster for future use
conif_r <- as(conif, "Raster")
# The name of this raster is undefined, so let's name it
names(conif_r) <- "conifer"

# # Testing it
plot(conif, main = "Conifer Land Cover and Wolf Points")
plot(wolf_sf, add=TRUE) # two points are off here but overall it looks fine


# Now let's pull in the elevation data
elevation <- rast(".\\Data\\Elevation2.tif")
# check the projection
crs(elevation) #26911
# reproject to match the others
elev2 <- elevation %>% terra::project("EPSG:4326", method = "near")
# convert to a raster for future use
elev2_r <- as(elev2, "Raster")

# # Testing how data looks
plot(elev2, main = "Elevation and Wolf Points")
plot(wolf_sf$geometry, add=TRUE) # This looks good now
```


```{r Set up Track Data}
# nest the data by the wolf ID
wolf_summer3 <- wolf_summer2 %>% nest(data = c(-wolfuid,-pack))
#str(wolf_summer3)
# test for duplicated time stamps
# would need to go through and do these by individual
# test <- wolf_summer3[duplicated(wolf_summer3$datetime_new),]

# testing making a track for one wolf
# wolf_track <- amt::make_track(tbl=wolf_summer3[[2]][[1]], .x=long, .y=lat, .t=datetime_new, crs = 4326) %>% amt::transform_coords(3857)


# trying it with a function
track_all <- wolf_summer3 %>%
 # make the track
 mutate(trk = map(data, function(d) {
   #transform from geographic to projected coord system 
    amt::make_track(tbl=d, .x=long, .y=lat, .t=datetime_new, crs = 4326) %>% # changed from tbl=d, .x=xcoord, .y=ycoord, .t=datetime_new
      amt::transform_coords(st_crs(conif))}))


#summarize sampling rates by individuals and then unnest 
samp_rates <- track_all %>% mutate(sr = lapply(trk, summarize_sampling_rate)) %>%
 dplyr::select(wolfuid, sr) %>% unnest(cols = c(sr))
samp_rates
# the highest median fix rate is 6 hours, so I'll be resampling them all to fit this definition
```

### Notes from class on extracting covariates
If covariates are extracted at the end of the step, they are typically included in the model as main effects, to answer questions of the type: How do covariates influence where the animal moves? 

In contrary, if covariates are extracted at the beginning of the step, they are typically included in the model as an interaction with movement characteristics (step length, log of the step length, or the cosine of the turn angle), to test hypotheses of the type: Do animals move faster/more directed, if they start in a given habitat? 

Finally, covariate values at the start and the end of a step can also be included in the model as an interaction with each other, to test hypotheses of the type: Are animals more likely to stay in a given habitat, if they are already in that habitat?
```{r}
# create an SSF dataframe
# If time:
# # you could also loop through this to see how it changes- recommended for multi scale nature of SSFs

track_resamp_all <- track_all %>%
   mutate(steps = map(trk, function(x) {
     x %>% amt::track_resample(rate = minutes(360), tolerance = minutes(10)) %>% # resample the tracks for each individual
      amt::filter_min_n_burst() %>% # filter them into bursts
       amt::steps_by_burst() %>% amt::random_steps() %>% # filter them into bursts again
       amt::extract_covariates(covariates = elev2, where = "both") %>% # this does both start and end | names the column the same as the column and adds on where 
       mutate(unique_step = paste(burst_,step_id_,sep="_")) %>% 
#       mutate(elev_start = factor(Elevation2_start)) %>% 
#       mutate(elev_end = factor(Elevation2_end)) %>% # filter them into bursts again
       amt::extract_covariates(covariates = conif, where = "both") %>% # this does both start and end | names the column the same as the column and adds on where 
       mutate(conifer_start = factor(HABITATTYPE_start)) %>% 
       mutate(conifer_end = factor(HABITATTYPE_end)) 
     }))

# Unnest this data
tracks_unnest <- track_resamp_all %>% 
  dplyr::select(wolfuid,pack, steps) %>% 
  unnest()

# Clean up the data
tracks_unnest <- tracks_unnest %>% rename(elev_start = Elevation2_start, elev_end = Elevation2_end) %>% select(-c(HABITATTYPE_start,HABITATTYPE_end))

tracks_unnest <- tracks_unnest %>% mutate(pack_fac=factor(pack))
str(tracks_unnest)
#tracks_unnest$pack <- as.factor(pack)
```

# Part 2: Statistical Analysis 
Negative coefficient: avoidance-ish, positive coefficient: selection

X as many times likely to be selected for
Difference between RSF and SSF is that SSF generates available points based on the movement process

## Exploratory analysis of step lengths and turning angles

Next, we make some summary plots of step length, turning angle as a function of day/night, and wet and other landcover types. We bundle them together using the `cowplot` package. 

```{r Look at Relationships}
# Look at step length and turning angle distributions
## sl_ for step length and ta_ for turning angle, time end is t2_

# for plotting:
tracks_plotting <- tracks_unnest %>% filter(is.na(conifer_end)==FALSE) %>% filter(is.na(elev_end)==FALSE)

tracks_used <- tracks_plotting %>% filter(case_ == TRUE)
tracks_avail <- tracks_plotting %>% filter(case_ == FALSE)
# separate by used and available 

# make a bar plot of sl_ for conifer_end = TRUE and FALSE
ggplot(data = tracks_plotting, mapping = aes(x = conifer_end, y = sl_)) + 
  geom_boxplot() + facet_wrap(~case_)
# doesn't look like there's a difference

ggplot(data = tracks_plotting, mapping = aes(x = elev_end, y = sl_)) + 
  geom_point()+ facet_wrap(~case_)
# slight increase at the end but no clear trend

# Test for interactions with step length with covariates
ggplot(data = tracks_plotting, mapping = aes(x = elev_end, y = sl_, color = conifer_end)) + 
  geom_point()+ facet_wrap(~case_)
# separate by used and available 

# we also don't really see a trend here
```


 Include step length * covariates in models 
 
```{r}
# Fit a naive GLM
nGLM_1 <- glm(case_~ I(conifer_end)+elev_end, data=tracks_unnest,family=binomial(link="logit"))
nGLM_1
# negative effect of elevation and conifer on selection

# Interpreting coefficients: wolves are slightly less likely (-0.03901) to use areas that are conifer than areas that aren't conifer
# taking a long time to run?
# nGLM_2 <- glm(case_~ elev_end, data=tracks_unnest,family=binomial(link="logit"))
# nGLM_2

# weak negative effect of conifer and elevation
# could be because wolves aren't using higher elevations in the summer?
```


# Fitting iSSF
Focus just on one season, and just 1 level of random effects at the pack level including intercepts and coefficients as recommended by Muff et al. (2020).  Finally, compare your ssf model estimated with a naive clogit, coxme, and glmmTMB. 

*Could try two stage*

```{r Simple SSF}
# hypothesis that habitat selection varies as a function of conifer presence, elevation, and the interaction of the conifer and elevation
# we didn't see any evidence in our exploratory analysis that the movement rate varies with any of our covariates
ssf_mod <- tracks_unnest %>% amt::fit_issf(case_ ~ conifer_end + elev_end + conifer_end:elev_end + strata(unique_step))

# took out random effects: + (1|pack_fac) + (0+pack_fac) 

summary(ssf_mod)

ssf_sum <- summary(ssf_mod$model)$coefficients

# Look at the outputs 
ssf_sum
AIC(ssf_mod)
```

If exponentiated coefficient is 1, the effect is super small


```{r Naive Clogit}
#no random effects
# This is taking a long time to run?
clogit1 <- clogit(case_ ~ conifer_end + elev_end + conifer_end:elev_end, data = tracks_unnest)
# clogit1 <- clogit(case_ ~ conifer_end + elev_end + conifer_end:elev_end, data = tracks_unnest)
#clogit1 <- clogit(case_ ~ I(landuseName=="Developed Open") + I(landuseName=="Developed Other") +I(landuseName=="Natural")+I(landuseName=="Crops") + strata(stratum), data = fisher6)
summary_clogit <- summary(clogit1)
coef(clogit1)
# line 802 in notes

```

```{r coxme model}
# random intercepts and random slopes 

clogitM1<- coxme(Surv(dt_,case_) ~ conifer_end + elev_end + conifer_end:elev_end +(1|pack_fac) + (0+pack_fac) , data=tracks_unnest)
AIC(clogitM1)
summary(clogitM1)
# line 838 in notes

```

```{r glmmTMB}
# look in previous notes for this
# Fit models to data
# making a random 
# land use interaction, stratum for step ID, random coefficient for each individual ID
TMBm1 <- glmmTMB(case_~ I(conifer_end) + elev_end + (1|pack) + (0 + pack), family = poisson, data = tracks_unnest, doFit = FALSE)

# change to pack
# 1|step_id_ is random intercept
# (0 + I(conifer)|wolf) is interaction between the two

TMBm1$parameters$theta[1] <-log(1e3) # set theta one (the 0 in the 0 + I(landuse_end)|id)

TMBm1$mapArg <-list(theta=factor(c(NA, 1:20))) # 15 random coefficients; four land cover types and six fishers (1 land cover type and 5 packs)
# tried 1-20 and none of them worked
# UNK use of these numbers 
glmm.TMB.random <- glmmTMB::fitTMB(TMBm1) # fit your model (from above) using these arguments
# ERROR: a map factor length must equal parameter length ##############################
summary(glmm.TMB.random)
# line 897 from class notes
coefTMB <- fixef(glmm.TMB.random)


v5 <- coefTMB$cond[2:5]
coefSum2 <- as.data.frame(cbind(coefSum, v5))
names(coefSum2) <- c("Naive", "clogit", "coxme", "two-stage iSSF", "glmmTMB")
coefSum2


# Run AIC on models
```

# Part 3: Inference




```{r Code Graveyard}

# resample this to six hours
# track_resamp_conif <- track_all %>%
#    mutate(steps = map(trk, function(x) {
#      x %>% amt::track_resample(rate = minutes(360), tolerance = minutes(10)) %>% # resample the tracks for each individual
#       amt::filter_min_n_burst() %>% # filter them into bursts
#        amt::steps_by_burst() %>% amt::random_steps() %>% # filter them into bursts again
#        amt::extract_covariates(covariates = conif, where = "both") %>% # this does both start and end | names the column the same as the column and adds on where 
#        mutate(unique_step = paste(burst_,step_id_,sep="_")) %>% 
#        mutate(conifer_start = factor(HABITATTYPE_start)) %>% 
#        mutate(conifer_end = factor(HABITATTYPE_end))
#      }))
# warnings here means that there are points off the map, there isn't too many of these so we're going to ignore these for this analysis



# create an SSF dataframe with elevation
# resample this to six hours
# track_resamp_elev <- track_all %>%
#    mutate(steps = map(trk, function(x) {
#      x %>% amt::track_resample(rate = minutes(360), tolerance = minutes(10)) %>% # resample the tracks for each individual
#       amt::filter_min_n_burst() %>% # filter them into bursts
#        amt::steps_by_burst() %>% amt::random_steps() %>% # filter them into bursts again
#        amt::extract_covariates(covariates = elev2, where = "both") %>% # this does both start and end | names the column the same as the column and adds on where 
#        mutate(unique_step = paste(burst_,step_id_,sep="_")) %>% 
#        mutate(elev_start = factor(Elevation2_start)) %>% 
#        mutate(elev_end = factor(Elevation2_end))
#      }))
# use start for SSF, use end for examining the relationship with covariates
# first specify the fix rate to resample it to (2 hours), set a tolerance of how much fudge factor to include around the 10 minutes
# stps <- amt::track_resample(track_all, rate = minutes(360), tolerance = minutes(10)) %>%
#   # define minimum number of locations in a burst
#   amt::filter_min_n_burst(min_n = 3) %>% amt::steps_by_burst() %>%
#   amt::time_of_day(include.crepuscule = FALSE)


# check the structure here
# str(stps, width = 80, strict.width = "no", nchar.max = 80, give.attr = FALSE)

# track_resampled_conif %>%
#   dplyr::select(id, steps) %>%
#   unnest()
# track_resampled_elev %>%
#   dplyr::select(id, steps) %>%
#   unnest()


# first filter out one individual from the data to make the track 
# wolf_ids <- unique(wolf_summer$wolfid)
# # loop through creating a track for each wolf ID
# for (id in wolf_ids){
#   #create a dataframe for each 
#   print(id)
#   individ_data <- wolf_summer %>% filter(wolfid == id)
#   # make a track data for each individual
#   print(nrow(individ_data))
# }

# Generate random steps
#Line 369

# establish to sample 9 available points
# this is defaulting to sampling at the end 
# m1 <-stps %>% amt::random_steps(n = 9) %>% # throwing an error
#   amt::extract_covariates(conif) %>%
#   mutate(conif = if_else(conif == "TRUE", 1, 0)) %>% 
#   amt::time_of_day(include.crepuscule = FALSE) %>%
#   mutate(unique_step = paste(burst_,step_id_,sep="_")) %>% 
#   mutate(log_sl_ = log(sl_)) -> d1

# Error in fitdistrplus::fitdist(x, "gamma", keepdata = FALSE, start = list(scale = scale_closed,  : 
#   data must be a numeric vector of length greater than 1

#Error in fitdistrplus::fitdist(x, "gamma", keepdata = FALSE, start = list(scale = scale_closed,  : 
#  data must be a numeric vector of length greater than 1


# set up the track data
#wolf_tracks <- wolf_summer3 %>% dplyr::select(x = `x_coord1`, y = `y_coord1`,
#           t = `time`)

# make a datetime 
#wolf_summer3 %>% separate(date, into = c(Day, Month, Year), sep="/")
#fake_data <- fake_data %>% mutate(Day = as.numeric(Day),Month=as.numeric(Month),Year=as.numeric(Year))
#                     
#fake_data <- fake_data %>% mutate(DateTime = make_datetime(year=Year,month=Month,day=Day,hour=Hour,min=Min,sec=Sec))

# make the date and time into a time
#wolf_summer4 <- wolf_summer %>% mutate(date = as.Date(date,format="%m/%d/%Y"))

#
# wolf_summer %>% mutate(time)
# DailyHours <- DailyHours %>% mutate(Date=strptime(Date, format="%Y%M%D"))

#prey_lancov <- rast(".\\Data\\lab6Stack.tif")
#plot(prey_lancov)
# extract covariates of interest
# elev <- prey_lancov[[7]]
# plot(elev)
# habtype <- prey_lancov[[10]]



# # Unite date and time into a column
# wolf_summer2 <- wolf_summer2 %>% unite(datetime,c(date,time),sep=" ")
# #  class(wolf_summer4$datetime) this is a character
# 
# # Reformat the date and time column into class datetime 
# wolf_summer2 <- wolf_summer2 %>% mutate(datetime_new=as.POSIXct(datetime,format="%m/%d/%y %H:%M"))
# # format that you specify for what it's coming in as
# 
# # remove the NAs in the datetime_new column
# wolf_summer2<-wolf_summer2[complete.cases(wolf_summer2[24]),]
# 
# 
# # # nest the data by the wolf ID
# # wolf_summer2 <- wolf_summer %>% nest(data=wolfid)
# # 
# # # make a track from this data
# wolf_track <- amt::make_track(wolf_summer2, x_coord1, y_coord1, datetime_new, crs = 4326)
# 
# # Make a track for multiple animals ######################
# # Line 558 from lab 10
# #unique(wolf1$pack)
# # select only the data you want
# # wolf_summer2 %>%
# #   dplyr::select(x = x_coord1, y = y_coord1,
# #               t = `datetime_new`, id = `wolfid`) # can add to filter out specific animals
# 
# # nest the data by the wolf ID
# wolf_summer3 <- wolf_summer2 %>% nest(data=wolfid)
# # 
# # wolves_track <- wolf_summer3 %>%
# #  # make the track
# #   mutate(trk = map(data, function(d) {
# #      amt::make_track(d, x_coord1, y_coord1, datetime_new, crs = 4326) }))
# # 
# # # %>%
# # #        amt::transform_coords(st_crs(elev2))
# 
# 
# # # make a track from this data
# wolf_track <- amt::make_track(wolf_summer3, x_coord1, y_coord1, datetime_new, crs = 4326)


# wolf_track_test <- as.numeric(wolf_track)

# check the fix rate - may need to resample to account for low fixes
# amt::summarize_sampling_rate(wolf_track)
# mean is 4.77 minutes 
# max is 192 minutes (1.5 hours)
# mean is 74 min, max is 390479 and min is zero
# this output is all in minutes 
# how many locations for each individual wolf?

# explore_1 <- tracks_unnest %>% dplyr::select(conifer_end, tod = t2_, sl_, ta_) %>%
#   gather(key, val, -conifer_end, -tod) %>%
#   filter(key == "sl_") %>%
#   ggplot(., aes(val, group = tod, fill = tod)) + geom_density(alpha = 0.5) +
#   #facet_wrap(~ landuse, nrow = 2) +
#   xlab("Step length [m]") + theme_light() +
#   ylab("Density") +
#   theme(legend.title = element_blank())
